---
title: "The impact of national health insurance on health services utilization in Lao People's Democratic Republic (Lao PDR)"
author: "Sekeun Daniel Yu (yus109@mcmaster.ca), Michel Grignon, Godefroy Emmanuel Guindon, Jean-Éric Tarride"
date: "September/2026"
---

# Setup
<br>
```{r setup}

# Set working directory
rm(list = ls(all.names = TRUE))
#install.packages("rstudioapi")
library(rstudioapi)
setwd(dirname(rstudioapi::getSourceEditorContext()$path))
getwd()

# Install and load packages
source("lao_nhi_0_package.r", echo = TRUE)

# Create directories
if(!dir.exists("./log")) {dir.create("./log")}

# Log
logpath <- file.path(getwd(), "log", "lao_nhi_3_baseline_and_overlap.log")
logrecord <- log_open(logpath, autolog = TRUE, show_notes = FALSE)
log_code()

```
<br>

# Load data
<br>
```{r load}

data.cln <- read_excel("./data export/02_clean_8_cln_q_dl.xlsx") # district
warnings()
str(data.cln)

data <- data.cln

# Check number of district
data %>% 
  filter(faclevel != "PH") %>% 
  summarise(count = n_distinct(district)) %>% # total district; 148
  put()

data %>% 
  filter(faclevel != "PH", treat == 1) %>% 
  summarise(count = n_distinct(district)) %>% # total district with NHI; 139
  put()

data %>% 
  group_by(province) %>% 
  filter(faclevel != "PH") %>% 
  summarise(count = n_distinct(district)) %>%
  put()

data %>% 
  group_by(province, district) %>% 
  summarize(n = n()) %>% 
  print(n = Inf) %>%
  put()

# Filter data
data <- data %>% 
  filter(
    !(province %in% c("15 Xekong")), # Drop NHI pilot province
    !(district %in% c("1008 Met", "1009 Viangkham", "1011 Mun",
                      "1306 Nong", "1408 Samouay")), # Drop NHI pilot districts
    !(district %in% c("0201 Phongsali")), # Drop district without CU5 in Q4/2015
    !str_detect(faclevel, "PH"), # Remove provincial hospitals
    !(quarter %in% c("2015.1", "2015.2", "2015.3"))) # Quarters to drop
 
# District with PH or HC only
# PH: XX00
# PSL 0205 Bounnua: HC only
# LNT 0301 Namtha: HC only
# ODX 0401 Xai: HC only
# HPN 0701 Xamnua: HC only
# XYB 0801 Xainyabouli: HC only
# XKG 0901 Pek: HC only
# BKX 1101 Pakxan: HC only
# KMN 1201 Thakhek: HC only
# SVK 1301 Kaysone Phomvihane: HC only
# SRV 1401 Salavan: HC only 
# SEK 1501 Lamam: HC only
# CPS 1601 Pakxe: HC only 
# ATP 1702 Samakhixai: HC only 
# XSB 1801 Anouvong: HC only 

# 1210 Khounkham - excluded in facility-level analysis but included in district-level analyses

# Excluding these 12 districts with HC only (due to the presence of PH), increases the DID estimates by about 0.05 points.
#data <- subset(data,District!="0205 Bounnua") # HC only
#data <- subset(data,District!="0301 Namtha") # HC only
#data <- subset(data,District!="0401 Xai") # HC only
#data <- subset(data,District!="0701 Xamnua") # HC only
#data <- subset(data,District!="0801 Xainyabouli") # HC only
#data <- subset(data,District!="0901 Pek") # HC only
#data <- subset(data,District!="1201 Thakhek") # HC only
#data <- subset(data,District!="1301 Kaysone Phomvihane") # HC only
#data <- subset(data,District!="1401 Salavan") # HC only
#data <- subset(data,District!="1601 Pakxe") # HC only
#data <- subset(data,District!="1702 Samakhixai") # HC only
#data <- subset(data,District!="1801 Anouvong") # HC only

# Create directories
if(!dir.exists("./output")) {dir.create("./output")}
if(!dir.exists("./graph")) {dir.create("./graph")}

```
<br>

# Table 1. Baseline characteristics
<br>
```{r tbl_1, warning=FALSE}

# package tableone
# http://www.danieldsjoberg.com/gtsummary-weill-cornell-presentation/#14
        
colnames(data)
str(data)

# District-level population characteristics
# Intervention  Control   
# 4, 5, 6, 8    9
table1a <- data %>%
 filter(quarter == 2016.2) %>%                    # filter unique quarter
 filter(group != 0) %>%                           # remove never treated
 dplyr::select(group, facnum, semp, urba, imps, litr, 
        opo5r, opu5r, ipo5r, ipu5r, iplos, delr) %>%
 mutate(across(c(delr, semp, urba, imps, litr), ~ .x * 100)) %>%
 mutate(treat_g = factor(ifelse(group == 9, "Control", "Intervention"),
                         levels = c("Intervention", "Control"))) %>%
 tbl_strata(strata = treat_g,
            ~tbl_summary(.x,
                         by = group,
                         statistic = list(
                           all_continuous() ~ "{mean} ({sd})"),
                         label = list(
                           facnum ~ "Mean number of health facilities",
                           semp ~ "Self-employed population, %",
                           urba ~ "Urban population, %",
                           imps ~ "Improved sanitation, % ",
                           litr ~ "Literacy among aged 15-64, %",
                           opo5r ~ "Outpatient visits ≥ 5, per 1,000",
                           opu5r ~ "Outpatient visits < 5, per 1,000",
                           ipo5r ~ "Inpatient admissions ≥ 5, per 100,000",
                           ipu5r ~ "Inpatient admissions < 5, per 100,000",
                           iplos ~ "Inpatient length of stay, days",
                           delr ~ "Institutional delivery, %"),
                         digits = list(
                           facnum ~ 0, semp ~ 1, urba ~ 1, imps ~ 1, litr ~ 1, 
                           opo5r ~ 0, opu5r ~ 0, ipo5r ~ 0, ipu5r ~ 0, 
                           iplos ~ 1,   # 1201 Thakhek missing
                           delr ~ 1),
                         missing = "no") %>%
              add_p(test = all_continuous() ~ "oneway.test",
                    test.args = list(var.equal = TRUE)) %>%
              modify_header(all_stat_cols() ~ "**{level}**<br>(n={n})",
                            p.value ~ "**p-value**") %>%
              remove_footnote_header(columns = everything()))  # remove footnote 1
# modify_column_alignment(columns = everything(), align = "center")

table1a
show_header_names(table1a) %>% put()

# imps: Percentage of population using improved sanitation
# litr: Literacy rate 15-64 old
# semp: Self-employment rate
# urba: Percentage of urban population

pnumdnum <- data %>% 
  group_by(group) %>%
  summarize(pcount = n_distinct(province),
            dcount = n_distinct(district)) %>%
  mutate(pcount = as.numeric(pcount),
         dcount = as.numeric(dcount)) %>%
  put()

pdcount <- data %>% 
  group_by(group) %>%
  summarize(pcount = n_distinct(province),
            dcount = n_distinct(district)) %>%
  mutate(pcount = as.numeric(pcount),
         dcount = as.numeric(dcount),
         p_by_d = pcount / dcount) %>%
  put()

table1 <- data %>%
 filter(quarter == 2016.2) %>%  # filter unique quarter
 filter(group != 0) %>%        # remove never treated
 mutate(facnum2 = facnum) %>%
 left_join(pdcount, by = c("group")) %>%
 dplyr::select(p_by_d, district, group, facnum, facnum2, 
        semp, urba, imps, litr, 
        opo5r, opu5r, ipo5r, ipu5r, iplos, delr) %>%
 mutate(district = 1) %>%
 mutate(across(c(delr, semp, urba, imps, litr), ~ .x * 100))

str(table1) %>% 
  put()

# Control group
table1c <- table1 %>%
 filter(group %in% c(5, 6, 8, 9)) %>%  # remove never treated
 tbl_summary(by = group,
             statistic = list(
               p_by_d ~ "{sum}",
               district ~ "{sum}",
               facnum ~ "{sum}",
               facnum2 ~ "{mean} ({sd})",
               semp ~ "{mean} ({sd})",
               urba ~ "{mean} ({sd})",
               imps ~ "{mean} ({sd})",
               litr ~ "{mean} ({sd})",
               opo5r ~ "{mean} ({sd})",
               opu5r ~ "{mean} ({sd})",
               ipo5r ~ "{mean} ({sd})",
               ipu5r ~ "{mean} ({sd})",
               iplos ~ "{mean} ({sd})",
               delr ~ "{mean} ({sd})"),
             label = list(
               p_by_d ~ "Province, n",
               district ~ "District, n",
               facnum ~ "Health facilities, n",
               facnum2 ~ "Mean number of health facilities",
               semp ~ "Self-employed population, %",
               urba ~ "Urban population, %",
               imps ~ "Access to improved sanitation, % ",
               litr ~ "Literacy rate among aged 15-64, %",
               opo5r ~ "Outpatient visits ≥ 5, per 1,000",
               opu5r ~ "Outpatient visits < 5, per 1,000",
               ipo5r ~ "Inpatient admissions ≥ 5, per 100,000",
               ipu5r ~ "Inpatient admissions < 5, per 100,000",
               iplos ~ "Inpatient length of stay, days",
               delr ~ "Institutional delivery, %"),
             digits = list(
               p_by_d ~ 0, district ~ 0, facnum ~ 0, facnum2 ~ 1,
               semp ~ 1, urba ~ 1, imps ~ 1, litr ~ 1,
               opo5r ~ 0, opu5r ~ 0, ipo5r ~ 0, ipu5r ~ 0,
               iplos ~ 1,   # 1201 Thakhek missing
               delr ~ 1),
             type = list(p_by_d ~ 'continuous',
                         district ~ 'continuous',
                         facnum ~ "continuous",
                         facnum2 ~ "continuous"),
             missing = "no") %>%
#                      all_continuous() ~ 1)) %>%
 modify_header(label = "**Variables**",
               stat_1 ~ "G5\n(q4/2016)\nn={n}",
               stat_2 ~ "G6\n(q1/2017)\nn={n}",
               stat_3 ~ "G8\n(q3/2017)\nn={n}",
               stat_4 ~ "G9\n(q4/2017)\nn={n}") %>%
 add_overall(last = TRUE, col_label = "All control\ngroups\nn={n}") %>%
 remove_footnote_header(columns = everything())  # remove footnote 1
# modify_column_alignment(columns = everything(), align = "center")
table1c %>% put()
show_header_names(table1c) %>% # province number not correct!
  put()

# Treated group
table1t <- table1 %>%
 filter(group != 9) %>%  # remove never treated
 tbl_summary(by = group,
             statistic = list(
               p_by_d ~ "{sum}",
               district ~ "{sum}",
               facnum ~ "{sum}",
               facnum2 ~ "{mean} ({sd})",
               semp ~ "{mean} ({sd})",
               urba ~ "{mean} ({sd})",
               imps ~ "{mean} ({sd})",
               litr ~ "{mean} ({sd})",
               opo5r ~ "{mean} ({sd})",
               opu5r ~ "{mean} ({sd})",
               ipo5r ~ "{mean} ({sd})",
               ipu5r ~ "{mean} ({sd})",
               iplos ~ "{mean} ({sd})",
               delr ~ "{mean} ({sd})"),
             label = list(
               p_by_d ~ "Province, n",
               district ~ "District, n",
               facnum ~ "Health facilities, n",
               facnum2 ~ "Mean number of health facilities",
               semp ~ "Self-employed population, %",
               urba ~ "Urban population, %",
               imps ~ "Access to improved sanitation, % ",
               litr ~ "Literacy rate among aged 15-64, %",
               opo5r ~ "Outpatient visits ≥ 5, per 1,000",
               opu5r ~ "Outpatient visits < 5, per 1,000",
               ipo5r ~ "Inpatient admissions ≥ 5, per 100,000",
               ipu5r ~ "Inpatient admissions < 5, per 100,000",
               iplos ~ "Inpatient length of stay, days",
               delr ~ "Institutional delivery, %"),
             digits = list(
               p_by_d ~ 0, district ~ 0, facnum ~ 0, facnum2 ~ 1,
               semp ~ 1, urba ~ 1, imps ~ 1, litr ~ 1,
               opo5r ~ 0, opu5r ~ 0, ipo5r ~ 0, ipu5r ~ 0,
               iplos ~ 1,   # 1201 Thakhek missing
               delr ~ 1),
             type = list(p_by_d ~ 'continuous',
                         district ~ 'continuous',
                         facnum ~ "continuous",
                         facnum2 ~ "continuous"),
             missing = "no") %>%
 modify_header(label = "**Variables**",
               stat_1 ~ "G4\n(q3/2016)\nn={n}",
               stat_2 ~ "G5\n(q4/2016)\nn={n}",
               stat_3 ~ "G6\n(q1/2017)\nn={n}",
               stat_4 ~ "G8\n(q3/2017)\nn={n}") %>%
 add_overall(last = TRUE, col_label = "All treated\ngroups\nn={n}") %>%
# modify_header(all_stat_cols() ~ "**{level}**\n(N={n})") %>%
 remove_footnote_header(columns = everything())  # remove footnote 1
# modify_table_body(~.x %>% dplyr::relocate(stat_0, .after = stat_4))
# modify_column_alignment(columns = everything(), align = "center")
table1t %>% put()
show_header_names(table1t) %>% # province number not correct!
  put()

tbl_merge <- 
  tbl_merge(
    list(table1t, table1c),
    tab_spanner = c("**Intervention**", "**Control**")) %>%
  modify_caption("Table 1. Characteristics of samples, by group")
tbl_merge

set_flextable_defaults(font.family = "Times New Roman")
tbl_merge %>%
 as_flex_table() %>%
 save_as_docx(path = "output/tbl_1_sample_characteristics.docx",
              pr_section = prop_section(
                page_size = page_size(orient = "landscape")))

# Add top header: Mean value (SD)
# Note: n is the total number of district-level units. Group (G) indicates group

```
<br>

## Baseline differences in outcomes and covariates
<br>
```{r baseline dif, warning=FALSE}

# Data is by district
dist_baseline <- data %>%
  filter(event == -1) %>%
  filter(group != 0) %>%
  group_by(district) %>%
  summarise(
    facnum = mean(facnum, na.rm = TRUE),
    semp = mean(semp * 100, na.rm = TRUE),
    urba = mean(urba * 100, na.rm = TRUE),
    imps = mean(imps * 100, na.rm = TRUE),
    litr = mean(litr * 100, na.rm = TRUE),
    opo5 = mean(opo5, na.rm = TRUE),
    opu5 = mean(opu5, na.rm = TRUE),
    ipo5 = mean(ipo5, na.rm = TRUE),
    ipu5 = mean(ipu5, na.rm = TRUE),
    ipday = mean(ipday, na.rm = TRUE),
    del = mean(del, na.rm = TRUE))

dist_stats_long <- dist_baseline %>%
  summarise(
    across(
      c(facnum, semp, urba, imps, litr,
        opo5, opu5, ipo5, ipu5, ipday, del),
      list(
        mean = ~mean(.x, na.rm = TRUE),
        median = ~median(.x, na.rm = TRUE),
        sd   = ~sd(.x, na.rm = TRUE),
        cv   = ~sd(.x, na.rm = TRUE) / mean(.x, na.rm = TRUE),
        min  = ~min(.x, na.rm = TRUE),
        max  = ~max(.x, na.rm = TRUE)),
      .names = "{.col}_{.fn}")) %>%
  pivot_longer(
    cols = everything(),
    names_to = c("variable", "stat"),
    names_sep = "_",
    values_to = "value")

dist_stats_long <- dist_stats_long %>%
  mutate(
    value = case_when(
      stat == "cv" ~ round(value, 2),   # keep precision
      TRUE ~ round(value, 0)))

dist_stats_long <- dist_stats_long %>%
  mutate(
    variable = recode(variable,
      facnum = "Mean number of health facilities",
      semp = "Self-employed population, %",
      urba = "Urban population, %",
      imps = "Improved sanitation, %",
      litr = "Literacy among aged 15–64, %",
      opo5 = "Outpatient visits ≥ 5",
      opu5 = "Outpatient visits < 5",
      ipo5 = "Inpatient admissions ≥ 5",
      ipu5 = "Inpatient admissions < 5",
      ipday = "Inpatient length of stay, days",
      del = "Institutional delivery"),
    stat = recode(stat,
      mean = "Mean",
      median = "Median",
      sd = "Standard_deviation",
      cv = "Coefficient_of_variation",
      min = "Minimum",
      max = "Maximum"))

dist_stats_table <- dist_stats_long %>%
  pivot_wider(names_from = stat, values_from = value)

dist_stats_table

# mean is always larger than median; meaning long right tail

write_xlsx(dist_stats_table, "./output/tbl_01_district_stats.xlsx")

```
<br>

# B5. Covariate balance and treatment timing
<br>
```{r selection, warning=FALSE}

data.cln <- read_excel("./data export/02_clean_8_cln_q_dl.xlsx") # district
warnings()
str(data.cln)

data <- data.cln

# Drop pilot districts and PH
data <- data %>%
 filter(
   !(province %in% c("15 Xekong")),  # Drop NHI pilot province
   !(district %in% c("1008 Met", "1009 Viangkham", "1011 Mun",
                     "1306 Nong", "1408 Samouay")),  # Drop NHI pilot districts
   !(district %in% c("0201 Phongsali")),  # Drop district without CU5 in Q4/2015
   !str_detect(faclevel, "PH"),  # Remove provincial hospitals
   !(quarter %in% c("2015.1", "2015.2", "2015.3")))  # Quarters to drop

df_dist <- data %>%
  filter(group > 0)

# ------------------------------------------------------------------------------
# Joint F-tests for equality of means across groups/cohorts
# ------------------------------------------------------------------------------

vars <- c("facnum", "semp", "urba", "imps", "litr",
          "opo5r", "opu5r", "ipo5r", "ipu5r", "iplos", "delr")

joint_equality <- lapply(vars, function(v) {
  model <- lm(as.formula(paste(v, "~ factor(group)")), data = df_dist)
  pval <- anova(model)[["Pr(>F)"]][1]
  data.frame(variable = v, p_value = round(pval, 3))}) %>% 
  bind_rows()

# Interpretation:
# Rejection -> observable jointly predict timing
# Non-rejection -> no evidence of selection on these observable

print(joint_equality)
write_xlsx(joint_equality, "./output/tbl_b_13_test_joint_equality.xlsx")

# ------------------------------------------------------------------------------
# Cohort-specific tests for equality of means using later-treated (not-yet + last) groups
# ------------------------------------------------------------------------------

run_test <- function(data, treat_group, time_period) {
  data %>%
    filter(time == time_period) %>%
    filter(group >= treat_group) %>%   # treated + not-yet-treated
    mutate(treated = ifelse(group == treat_group, 1, 0))}

df_dist_4 <- run_test(df_dist, 4, 3)  # Q2 2016
df_dist_5 <- run_test(df_dist, 5, 4)  # Q3 2016
df_dist_6 <- run_test(df_dist, 6, 5)  # Q4 2016
df_dist_8 <- run_test(df_dist, 8, 7)  # Q2 2017

vars <- c("facnum", "semp", "urba", "imps", "litr",
          "opo5r", "opu5r", "ipo5r", "ipu5r", "iplos", "delr")

get_stats <- function(df) {
  res <- sapply(vars, function(v) {
    treated_vals <- df[df$treated == 1, ][[v]]
    control_vals <- df[df$treated == 0, ][[v]]
    mean_treated <- mean(treated_vals, na.rm = TRUE)
    mean_control <- mean(control_vals, na.rm = TRUE)
    pval <- t.test(treated_vals, control_vals)$p.value
    c(mean_treated, mean_control, pval)})
  t(res)}

res_4 <- get_stats(df_dist_4)
res_5 <- get_stats(df_dist_5)
res_6 <- get_stats(df_dist_6)
res_8 <- get_stats(df_dist_8)

results <- cbind(
  res_4[,1], res_4[,2], res_4[,3],
  res_5[,1], res_5[,2], res_5[,3],
  res_6[,1], res_6[,2], res_6[,3],
  res_8[,1], res_8[,2], res_8[,3])

results <- as.data.frame(results)
results$Variable <- vars

# Reorder columns
results <- results[, c("Variable", names(results)[1:12])]

# Define variable groups
covariates_pct <- c("semp", "urba", "imps", "litr")       # percentages -> ×100
covariates_raw <- c("facnum")                             # counts -> no scaling
outcomes0      <- c("opo5r", "opu5r", "ipo5r", "ipu5r")   # no decimals
outcomes1      <- c("iplos")                              # 1 decimal
outcomes2      <- c("delr")                               # percentages -> ×100


# Identify mean columns
mean_cols <- c(2,3,5,6,8,9,11,12)

# Apply formatting row by row
for (i in 1:nrow(results)) {
  var <- results$Variable[i]
  
  vals <- as.numeric(results[i, mean_cols])
  
  if (var %in% covariates_pct) {
    vals <- round(vals*100, 0)
  } else if (var %in% covariates_raw) {
    vals <- round(vals, 1)
  } else if (var %in% outcomes0) {
    vals <- round(vals, 0)
  } else if (var %in% outcomes1) {
    vals <- round(vals, 1)
  } else if (var %in% outcomes2) {
    vals <- round(vals*100, 0)}

  results[i, mean_cols] <- vals}

# Format p-values (3 decimals, no scientific notation)
format_p <- function(p) {
  ifelse(p < 0.001, "<0.001", sprintf("%.3f", as.numeric(p)))}

p_cols <- c(4, 7, 10, 13)
results[, p_cols] <- lapply(results[, p_cols], format_p)

colnames(results) <- c(
  "Variable",
  "G4 (treated)", "G5,6,8,9 (control)", "p-value (Q2/2016)",
  "G5 (treated)", "G6,8,9 (control)", "p-value (Q3/2016)",
  "G6 (treated)", "G8,9 (control)", "p-value (Q4/2016)",
  "G8 (treated)", "G9 (control)", "p-value (Q2/2017)")

results
write_xlsx(results, "./output/tbl_b_13_test_group_specific.xlsx")

# ------------------------------------------------------------------------------
# Joint F-test for selection on observables
# ------------------------------------------------------------------------------

df_dist <- data %>% 
  filter(group > 0 & quarter == "2016.2") %>%
  dplyr::select(group, province, district, pnum, dnum, event,
         facnum, semp, urba, imps, litr,
         opo5r, opu5r, ipo5r, ipu5r, iplos, delr)

# Run regression
timing_reg <- lm(group ~ facnum + semp + urba + imps + litr, 
                 data = df_dist)
summary(timing_reg)

coef_table <- tidy(timing_reg, conf.int = TRUE)
model_stats <- glance(timing_reg)

wb <- createWorkbook()
addWorksheet(wb, "Coefficients")
writeData(wb, "Coefficients", coef_table)
addWorksheet(wb, "Model Stats")
writeData(wb, "Model Stats", model_stats)
saveWorkbook(wb, "./output/tbl_b_14_test_treatment_timing.xlsx",
             overwrite = TRUE)

```
<br>

# Figure 1-A. Population coverage by social health protection schemes from 2008 to 2021
<br>
```{r fig_01_a, warning=FALSE}

data.shp <- read_excel("./data import/lao_shp_schemes.xlsx", 
                       sheet = "final") # district
data.shp <- data.shp %>% 
  pivot_longer(cols = -"scheme", names_to = "year", values_to = "population")

tpop <- data.shp %>%
  filter(scheme == "Total population") %>%
  dplyr::select(-scheme) %>%
  rename(total_population = population, year = year)

data.shp <- data.shp %>%
  filter(scheme != "Total population") %>%
  left_join(tpop, by = "year") %>%
  mutate(pct = population / total_population)

data.shp %>% distinct(scheme) %>% put()

scheme_list <- c("Formal sector Defense",
                 "Formal sector Civil servants",
                 "Formal sector Private employees",
                 "Informal sector Health Equity Fund for the poor (HEF)",
                 "Free mothers (FMAT)",
                 "Free children under 5 (FCU5)",
                 "Informal sector Community-based health insurance (CBHI)",
                 "Informal sector National Health Insurance (NHI)")
scheme_label <- c("Formal sector: Military and Police",
                  "Formal sector: Civil servants (LSSO-SASS)",
                  "Formal sector: Private enterprise workers (LSSO-SSO)", 
                  "Informal sector: Health Equity Fund for the poor (HEF)",
                  "Informal sector: Free maternity (FMAT)",
                  "Informal sector: Free children under five (FCU5)",
                  "Informal sector: Community-based health insurance (CBHI)",
                  "Informal sector: National Health Insurance (NHI)")

data.shp <- data.shp %>%
  mutate(scheme2 = factor(scheme, 
                          levels = scheme_list, 
                          labels = scheme_label))

write_xlsx(data.shp, "./output/fig_01_population_scheme.xlsx")

hiro <- met.brewer("Hiroshige") # Red -> Blue
mypalette <- hiro[c(1:5, 7:9)]
#hoku <- met.brewer("Hokusai3")[2]
#paqu <- met.brewer("Paquin")[7]
#paqu <- met.brewer("Paquin")[8]
vgh3 <- met.brewer("VanGogh3")[3]
mypalette[1] <- vgh3
tpopn <- 7337783

# plot -------------------------------------------------------------------------

plot_pop <- ggplot(
  data.shp, aes(fill = scheme2, y = population, x = year)) + 
  geom_bar(position = position_stack(reverse = TRUE), 
           color = "grey25", stat = "identity", lwd = 0.2) +
  scale_y_continuous(limits = c(0, tpopn), expand = c(0, 0), 
                     labels = scales::comma) +
  scale_fill_manual(values = rev(mypalette), breaks = scheme_label) +
  ylab("Population") + 
  xlab("Year") +
  annotate("text", x = 7.5, y = 2900000, label = "33%", size = 3.3,
           family = "Source Sans 3") +
  geom_segment(aes(x = 7.7, y = 2650000, xend = 8, yend = 2400000),
               arrow = arrow(type = "closed", length = unit(0.05, "in")),
               color = "grey30") +
  annotate("text", x = 9.5, y = 7150000, label = "93%", size = 3.3,
           family = "Source Sans 3") +
  geom_segment(aes(x = 9.7, y = 6900000, xend = 10, yend = 6650000),
               arrow = arrow(type = "closed", length = unit(0.05, "in")),
               color = "grey30") +
  ggtitle("(a) Population coverage by health financing scheme, 2008-2021") +
  theme_light() +
  theme(
    text = element_text(family = "Source Sans 3"),
    panel.border = element_blank(),  
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black"),
    plot.title.position = "plot",
    plot.title = element_markdown(hjust = 0.7,          # fully left align
                                  size = 11, colour = "black",
                                  margin = margin(b = 20)),
    plot.margin = margin(t = 0.1, r = 0.05, b = 0.1, l = 0.1, unit = "in"),
    axis.title.x = element_text(color = "black", size = 10.5, 
                                margin = margin(0.15, 0, 0, 0, "in")),
    axis.title.y = element_text(color = "black", vjust = 3, size = 10.5),
    axis.ticks = element_line(color = "black"),
    axis.text.y = element_text(color = "black", size = 10.5),
    axis.text.x = element_text(color = "black", size = 10, 
                               angle = 45, vjust = 0.75, hjust = 0.75),
    legend.position = 'bottom',
    legend.box = "vertical",          # put legend items in a vertical box
    legend.box.just = "top",
    legend.title = element_blank(),
    legend.key.size = unit(0.2, "in"),
    legend.key.spacing.y = unit(0.025, "in"),
    legend.text = element_text(size = 10)) + 
  guides(fill = guide_legend(ncol = 1, reverse = TRUE, byrow = TRUE))

# Vector file
pdf("./graph/fig_01_population_scheme.pdf", width = 6.3, height = 6.6)
print(plot_pop)
dev.off()

# plot 2 (y-axis unit million) -------------------------------------------------

tpopn <- 7.337783
data.shp_v <- data.shp %>%
  mutate(population = population / 1000000,
         total_population = total_population / 1000000)
  
plot_pop_v <- ggplot(
  data.shp_v, aes(fill = scheme2, y = population, x = year)) + 
  geom_bar(position = position_stack(reverse = TRUE), 
           color = "grey25", stat = "identity", lwd = 0.2) +
  scale_y_continuous(limits = c(0, tpopn), expand = c(0, 0), 
                     labels = scales::comma) +
  scale_fill_manual(values = rev(mypalette), breaks = scheme_label) +
  ylab("Population (million)") + 
  xlab("Year") +
  annotate("text", x = 7.5, y = 2.90, label = "33%", size = 3.3,
           family = "Source Sans 3") +
  geom_segment(aes(x = 7.7, y = 2.65, xend = 8, yend = 2.40),
               arrow = arrow(type = "closed", length = unit(0.05, "in")),
               color = "grey30") +
  annotate("text", x = 9.5, y = 7.15, label = "93%", size = 3.3,
           family = "Source Sans 3") +
  geom_segment(aes(x = 9.7, y = 6.90, xend = 10, yend = 6.65),
               arrow = arrow(type = "closed", length = unit(0.05, "in")),
               color = "grey30") +
  ggtitle("(a) Population coverage by health financing scheme, 2008-2021") +
  theme_light() +
  theme(
    text = element_text(family = "Source Sans 3"),
    panel.border = element_blank(),  
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black"),
    plot.title.position = "plot",
    plot.title = element_markdown(hjust = 0.7,          # fully left align
                                  size = 11, colour = "black",
                                  margin = margin(b = 20)),
    plot.margin = margin(t = 0.1, r = 0.05, b = 0.05, l = 0.2, unit = "in"),
    axis.title.x = element_text(color = "black", size = 10.5, 
                                margin = margin(0.15, 0, 0, 0, "in")),
    axis.title.y = element_text(color = "black", vjust = 3, size = 10.5),
    axis.ticks = element_line(color = "black"),
    axis.text.y = element_text(color = "black", size = 10.5),
    axis.text.x = element_text(color = "black", size = 10, 
                               angle = 45, vjust = 0.75, hjust = 0.75),
    legend.position = 'bottom',
    legend.box = "vertical",          # put legend items in a vertical box
    legend.box.just = "top",
    legend.title = element_blank(),
    legend.key.size = unit(0.2, "in"),
    legend.key.spacing.y = unit(0.025, "in"),
    legend.text = element_text(size = 10)) + 
  guides(fill = guide_legend(ncol = 1, reverse = TRUE, byrow = TRUE))

# Vector file
pdf("./graph/fig_01_population_scheme_v.pdf", width = 6.1, height = 6.6)
print(plot_pop_v)
dev.off()

```
<br>

# Figure 1-B. Timeline of NHI expansion from 2015 to 2019
<br>
```{r fig_01_b, warning=FALSE}

data.cln <- read_excel("./data export/02_clean_8_cln_q_dl.xlsx") # district

data.cln %>%
 filter(faclevel != "PH") %>%
 summarise(count = n_distinct(district)) %>%
 put()

data.cln %>%
 filter(faclevel != "PH") %>%
 filter(pnum != 1) %>%
 summarise(count = n_distinct(district)) %>%
 put()

data.num <- data.cln %>%
 filter(pnum != 1) %>%
 filter(faclevel != "PH") %>%
 put()

data.num <- data.num %>%
  dplyr::select(quarter, pnum, treat, group) %>%
  mutate(quarter = as.factor(quarter),
         group = as.factor(group))
unique(data.num$group) %>% put()
str(data.num) %>% put()

prov_label <- c("Q3 2016 (Attapeu, Luang Namtha, Xaysomboun)",
                "Q4 2016 (Salavanh)",
                "Q1 2017 (Bolikhamxay, Houaphanh, Xieng Khouang, Oudomxay, Phongsaly)",
                "Q2 2017 (Sekong)",
                "Q3 2017 (Savannakhet, Champasak)",
                "Q4 2017 (Bokeo, Khammouane, Luang Prabang, Sayabouly, Vientiane)")

data.num <- data.num %>%
  mutate(
    group = factor(group, labels = c("Q3 2016", "Q4 2016", "Q1 2017", 
                                     "Q2 2017", "Q3 2017", "Q4 2017")),
    quarter = factor(quarter, labels = c("Q4.2015", "Q1.2016", "Q2.2016", 
                                         "Q3.2016", "Q4.2016", "Q1.2017",
                                         "Q2.2017", "Q3.2017", "Q4.2017", 
                                         "Q1.2018", "Q2.2018", "Q3.2018", 
                                         "Q4.2018", "Q1.2019")),
    group2 = factor(group, labels = prov_label))

naqrt <- c("Q1.2015", "Q2.2015", "Q3.2015", "Q4.2015", "Q1.2016", "Q2.2016")
total <- data.num %>%
  group_by(quarter) %>%
  summarise(count = sum(treat)) %>%
  mutate(count = replace(count, quarter %in% naqrt, NA))

mypalette <- met.brewer("Hokusai3", 6, "discrete") # Paquin / Hiroshige
names(mypalette) <- prov_label

# Legend with province names (aggregated blocked), bar outline
data.num2 <- data.num %>%
  group_by(group2, quarter) %>%
  summarise(count = sum(treat)) %>%
  setNames(c("intv_group", "quarter", "count"))

write_xlsx(data.num2, "./output/fig_01_district_num.xlsx")

plot_district <- ggplot(
  data.num2, aes(fill = intv_group, y = count, x = quarter)) + 
  geom_bar(position = position_stack(reverse = TRUE), 
           color = "grey25", stat = "identity", lwd = 0.2) +
  geom_text(aes(x = quarter, y = count + 5, label = count, fill = NULL), 
            data = total, size = 3, fontface = "italic") +
  scale_y_continuous(limits = c(0, 155), expand = c(0, 0)) +
  scale_fill_manual(
    values = mypalette,
    breaks = prov_label) +
  labs(fill = NULL,
       y = "Number of districts", 
       x = "Quarter") +
  ggtitle("(b) Timeline of NHI expansion, 2015-2019") +
  theme_light() +
  theme(
    text = element_text(family = "Source Sans 3"), # Font style
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black"),
    plot.title.position = "plot",
    plot.title = element_markdown(
      hjust = 0.6,          # fully left align
      size = 11,
      colour = "black",
      margin = margin(b = 20)),
    plot.margin = margin(t = 0.1, r = 0.5, b = 0.1, l = 0.1, unit = "in"),
    axis.title.x = element_text(color = "black", size = 10.5, 
                                margin = margin(t = 0.15, unit = "in")),
    axis.title.y = element_text(color = "black", vjust = 2, size = 10.5),
    axis.ticks = element_line(color = "black"),
    axis.text.y = element_text(color = "black", size = 10.5),
    axis.text.x = element_text(color = "black", size = 10, 
                               angle = 45, vjust = 0.75, hjust = 0.75),
    # legend.title = element_text(color = "black", size = 10, 
    #                             margin = margin(b = 0.07, unit = "in")),
    legend.position = 'bottom',
      legend.box = "vertical",          # put legend items in a vertical box
  legend.box.just = "top",
    legend.key.size = unit(0.2, "in"),
    legend.key.spacing.y = unit(0.025, "in"),
    legend.text = element_text(size = 10.5)) +
guides(fill = guide_legend(ncol = 1, reverse = FALSE, byrow = TRUE))

# Vector file
pdf("./graph/fig_01_district_num.pdf", width = 5.5, height = 6)
print(plot_district)
dev.off()

# Combine graphs
plot_pop_clean  <- plot_pop  + theme(plot.margin = margin(r = 30))
plot_dist_clean <- plot_district + theme(
  legend.box.margin = margin(t = -35))
plot_pop_dist <- plot_pop_clean + plot_dist_clean + plot_layout(ncol = 2)

# Save
pdf("./graph/fig_01_pop_district.pdf", width = 10, height = 6.5)
print(plot_pop_dist)
dev.off()

```
<br>

# Figure 1-B. Choropleth map
<br>
```{r fig_01_b, warning=FALSE}

# Source
# https://data.humdata.org/dataset

# Label and color scheme
prov_label <- c("Q3 2016 (Attapeu, Luang Namtha, Xaysomboun)",
                "Q4 2016 (Salavanh)",
                "Q1 2017 (Bolikhamxay, Houaphanh, Xieng Khouang, Oudomxay, Phongsaly)",
                "Q2 2017 (Sekong)",
                "Q3 2017 (Savannakhet, Champasak)",
                "Q4 2017 (Bokeo, Khammouane, Luang Prabang, Sayabouly, Vientiane)")

hiro <- met.brewer("Hiroshige") # Red -> Blue
mypalette <- hiro[c(2, 2, 3, 5, 6, 7)]
mypalette[1] <- "#95c36e" # "#d8d97a"
mypalette[2] <- "#f78462" # "#d8d97a"
mypalette[3] <- "#fbc06a" # Hiroshige 3,4 between "#f7aa58" "#ffd06f"
mypalette[6] <- "#64aac9" # Hiroshige 7,8 between "#f7aa58" "#ffd06f", 37.5%

names(mypalette) <- prov_label
map_palette <- c(mypalette, "Never-treated province (Vientiane Capital)" = "grey85")

# Geospatial data
lao_admin2 <- st_read("data import/map_boundaries.shp/lao_admin2.shp")
thailand <- st_read("data import/map_boundaries.shp/tha_admin0.shp")
vietnam  <- st_read("data import/map_boundaries.shp/vnm_admin0.shp")
china  <- st_read("data import/map_boundaries.shp/chn_admbnda_adm0_ocha_2020.shp")
cambodia  <- st_read("data import/map_boundaries.shp/khm_admin0.shp")
myanmar  <- st_read("data import/map_boundaries.shp/mmr_admin0.shp")

# Match CRS to Laos
thailand <- st_transform(thailand, st_crs(lao_admin2))
vietnam  <- st_transform(vietnam,  st_crs(lao_admin2))
china  <- st_transform(china,  st_crs(lao_admin2))
cambodia  <- st_transform(cambodia,  st_crs(lao_admin2))
myanmar  <- st_transform(myanmar,  st_crs(lao_admin2))

sf_use_s2(FALSE)
vietnam <- st_make_valid(vietnam)
cambodia <- st_make_valid(cambodia)

thai_border <- st_union(thailand)
viet_border <- st_union(vietnam)
camb_border <- st_union(cambodia)
chin_border <- st_union(china)
myan_border <- st_union(myanmar)

viet_border_s <- st_simplify(viet_border, dTolerance = 0.01)

laos_border <- st_union(lao_admin2)
thai_border_clean <- st_difference(thai_border, laos_border)
viet_border_clean <- st_difference(viet_border_s, laos_border)
camb_border_clean <- st_difference(camb_border, laos_border)
chin_border_clean <- st_difference(chin_border, laos_border)
myan_border_clean <- st_difference(myan_border, laos_border)
sf_use_s2(TRUE)  # turn it back on

country_labels <- data.frame(
  country = c("Thailand", "Vietnam"),
  x = c(101.5, 105.5),  # longitude (x-axis)
  y = c(15.5, 21.7))    # latitude  (y-axis)

laos_bbox <- st_bbox(lao_admin2)
label_size <- 2

# Laos data
lao_map <- lao_admin2 %>%
  arrange(adm2_pcode) %>%
  mutate(
    group_num = case_when(
      adm1_pcode %in% c("LA17", "LA03", "LA18") ~ 4,
      adm1_pcode == "LA14" ~ 5,
      adm1_pcode %in% c("LA11", "LA07", "LA09", "LA04", "LA02") ~ 6,
      adm1_pcode == "LA15" ~ 7,
      adm1_pcode %in% c("LA13", "LA16") ~ 8,
      adm1_pcode %in% c("LA05", "LA12", "LA06", "LA08", "LA10") ~ 9,
      TRUE ~ NA_real_),
    group = case_when(
      is.na(group_num) ~ "Never-treated province (Vientiane Capital)",
      TRUE ~ prov_label[group_num - 3]),
    group = factor(
      group, levels = c(prov_label, "Never-treated province (Vientiane Capital)"))) %>%
  dplyr::select(-group_num)

# Plot
plot_map <- ggplot() +
  geom_sf(data = chin_border_clean, fill = "grey97", color = NA, linewidth = 0) +
  geom_sf(data = thai_border_clean, fill = "grey97", color = "grey70", linewidth = 0.2) +
  geom_sf(data = viet_border_clean, fill = "grey97", color = "grey70", linewidth = 0.2) +
  geom_sf(data = camb_border_clean, fill = "grey97", color = "grey70", linewidth = 0.2) +  
  geom_sf(data = myan_border_clean, fill = "grey97", color = "grey70", linewidth = 0.2) +
  geom_sf(data = lao_map, aes(fill = group), color = "black", linewidth = 0.2) +
  scale_fill_manual(values = map_palette, name = NULL, drop = FALSE) +
  coord_sf(xlim = c(laos_bbox["xmin"] - 0.2, laos_bbox["xmax"] + 0.2),
           ylim = c(laos_bbox["ymin"] - 0.2, laos_bbox["ymax"] + 0.2),
           expand = FALSE) +
  theme_light() +
  theme(
    text = element_text(family = "Source Sans 3"), # Font style
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.ticks.length = unit(0, "in"),
    axis.title = element_blank(),
    axis.line = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(colour = "grey75", fill = NA, linewidth = 0.1),
    legend.position = "none",
    plot.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0.1, unit = "in"))

plot_map

# Save
plot_map_alone <- plot_map +
  geom_text(data = country_labels, aes(x = x, y = y, label = country),
            color = "grey40", size = 3.5, family = "Source Sans 3", 
            fontface = "italic")
  
pdf("./graph/fig_01_map_alone.pdf", width = 5.3, height = 6.1)
print(plot_map_alone)
dev.off()

# Add legend and title
plot_map1 <- plot_map_alone + 
  ggtitle("(b) Timing of NHI introduction across provinces and districts, 2016-2017") +
  theme(
    plot.title = element_markdown(hjust = 0.6, # fully left align
                                  size = 11, colour = "black", 
                                  margin = margin(b = 20)),
    legend.position = "bottom",
    legend.box      = "vertical",
    legend.box.just = "top",
    legend.key.size = unit(0.2, "in"),
    legend.key.spacing.y = unit(0.025, "in"),
    legend.text     = element_text(size = 10),
    plot.margin = margin(t = 0.1, r = 1.1, b = 0, l = 1.5, unit = "in")) +
  guides(fill = guide_legend(ncol = 1, reverse = FALSE, byrow = TRUE))

# Standalone graph with legend and title
plot_map2 <- plot_map1 +
  theme(
    plot.margin = margin(t = 0, r = 0.2, b = 0, l = 0.2, unit = "in"),
    plot.title = element_markdown(hjust = 0.6, # fully left align
                                  size = 11, colour = "black",
                                  margin = margin(t = 1.5, b = 15))) 

pdf("./graph/fig_01_map.pdf", width = 5.4, height = 7.8)
print(plot_map2)
dev.off()

# Combine graphs (patchwork package) - horizontal
plot_pop_clean  <- plot_pop  + theme(plot.margin = margin(r = 27)) # legend.justification = "left"
plot_map_clean <- plot_map1 +
  theme(legend.box.margin = margin(t = -16))

plot_pop_map <- plot_pop_clean + plot_map_clean + plot_layout(ncol = 2) +
  inset_element(plot_map_alone, left = -0.5, right = 1.4, top = 1.05, bottom = -0.22)

# Save
pdf("./graph/fig_01_pop_map.pdf", width = 10.5, height = 6.4)
print(plot_pop_map)
dev.off()

# Combine graphs (patchwork package) - horizontal_v2 (reduced label) -> not complete
# plot_pop_clean  <- plot_pop_v  + theme(plot.margin = margin(r = 27)) # legend.justification = "left"
# plot_map_clean <- plot_map1 +
#   theme(legend.box.margin = margin(t = -16))
# 
# plot_pop_map <- plot_pop_clean + plot_map_clean + 
#   plot_layout(ncol = 2, widths = c(1, 1)) +
#   inset_element(plot_map_alone, left = -0.5, right = 1.4, top = 1.05, bottom = -0.22)
# 
# # Save
# pdf("./graph/fig_01_pop_map_h_v1.pdf", width = 10.5, height = 6.4)
# print(plot_pop_map)
# dev.off()
# 
# # Combine graphs (patchwork package) - vertical
# plot_pop_clean  <- plot_pop_v +
#   theme(plot.title = element_markdown(
#     hjust = 0.7, size = 11, colour = "black",
#     margin = margin(b = 0.165, unit = "in")),
#     legend.text = element_text(size = 10.5),
#     legend.box.margin = margin(t = -0.1, l = -1, unit = "in"),
#     plot.margin = margin(0, 0, 0.245, 0.01, unit = "in"))
# 
# plot_map_clean <- plot_map1 +
#   theme(plot.title = element_markdown(
#     hjust = 0.6, size = 11, colour = "black",
#     margin = margin(b = 0.165, l = -0.1, unit = "in")),
#     legend.text = element_text(size = 10.5),
#     legend.box.margin = margin(t = -0.07, l = -0.08, unit = "in"),
#     plot.margin = margin(0, 0, 0, 0, unit = "in"))
# 
# #plot_pop_map_v <- plot_pop_clean + plot_map_clean + plot_layout(nrow = 2)
# plot_pop_map_v <- plot_pop_clean / plot_map_clean + #v3
#   #plot_layout(heights = c(0.8, 1.2))
#   #plot_layout(heights = c(0.8, 1.27)) #v3
#   plot_layout(heights = c(0.8, 1.3)) #v1, v4
#   
# # Save
# #pdf("./graph/fig_01_pop_map_v.pdf", width = 11, height = 16)
# #pdf("./graph/fig_01_pop_map_v2.pdf", width = 5, height = 13.75)
# #pdf("./graph/fig_01_pop_map_v3.pdf", width = 5, height = 14.25)
# #pdf("./graph/fig_01_pop_map_v4.pdf", width = 5.1, height = 14.4)
# pdf("./graph/fig_01_pop_map_v1.pdf", width = 5.12, height = 14.35)
# print(plot_pop_map_v)
# dev.off()

# Combine graphs (cowplot package) - vertical
plot_pop_clean <- plot_pop_v +
  theme(plot.title = element_markdown(
    hjust = 0.7, size = 11, colour = "black", 
    margin = margin(b = 0.165, unit = "in")),
    legend.text = element_text(size = 10.5),
    legend.box.margin = margin(t = -0.1, l = -0.1, unit = "in"),
    plot.margin = margin(0, 0, 0.245, 0.01, unit = "in"))

plot_map_clean <- plot_map1 +
  theme(plot.title = element_markdown(
    hjust = 0.6, size = 11,
    colour = "black", margin = margin(b = 0.165, unit = "in")),
    legend.text = element_text(size = 10.5),
    legend.box.margin = margin(t = -0.07, l = 0.65, unit = "in"),
    plot.margin = margin(0, 0, 0, 0.01, unit = "in"))

plot_pop_map_v2 <- ggdraw() +
  theme(plot.background = element_rect(fill = "white", colour = NA)) +
  draw_plot(plot_pop_clean, x = 0.02, y = 0.545, 
            width = 0.97, height = 0.45) +
  draw_plot(plot_map_clean, x = 0.02, y = 0.00,
            width = 0.96, height = 0.545)

pdf("./graph/fig_01_pop_map_v2.pdf", 
    width = 5.12, height = 14.35)
print(plot_pop_map_v2)
dev.off()


```
<br>

# Figure 2. Analysis Framework
<br>
```{r fig_02, warning=FALSE}

fig_font <- "Source Sans 3"

inner_line <- 0.275
outer_line <- 0.4

# Column positions
x0 <- 0.5

w_type <- 1.75
w_prov <- 1.75
w_dist <- 1.75
w_group <- 2.8
w_time <- 1.8

x_type_min <- x0
x_type_max <- x_type_min + w_type
x_type_mid <- mean(c(x_type_min, x_type_max))

x_int_min <- x_type_min
x_int_max <- x_type_mid

x_con_min <- x_type_mid
x_con_max <- x_type_max

x_prov_min <- x_type_max
x_prov_max <- x_prov_min + w_prov

x_dist_min <- x_prov_max
x_dist_max <- x_dist_min + w_dist

x_group_min <- x_dist_max
x_group_max <- x_group_min + w_group

time_start <- x_group_max

time_x <- tibble(
  t = 1:9,
  xmin = time_start + (0:8) * w_time,
  xmax = time_start + (1:9) * w_time,
  xmid = (xmin + xmax) / 2)

x_end <- max(time_x$xmax)

# Row positions, controlled by row_h
row_h <- 0.65

y_body_top <- 5.5

y_rows <- tibble(
  row = 1:5,
  ymax = y_body_top - (0:4) * row_h,
  ymin = ymax - row_h,
  ymid = (ymin + ymax) / 2)

y_t_min <- y_body_top
y_t_max <- y_t_min + row_h
y_t_mid <- mean(c(y_t_min, y_t_max))

y_q_min <- y_t_max
y_q_max <- y_q_min + row_h
y_q_mid <- mean(c(y_q_min, y_q_max))

y_year_min <- y_q_max
y_year_max <- y_year_min + row_h
y_year_mid <- mean(c(y_year_min, y_year_max))

y_bottom <- min(y_rows$ymin)
y_top <- y_year_max

# Data
group_df <- tibble(
  group = c(4, 5, 6, 8, 9),
  province = c(3, 1, 5, 2, 5),
  district = c(15, 7, 37, 24, 46),
  ymid = y_rows$ymid)

status_df <- crossing(
  group = c(4, 5, 6, 8, 9),
  t = 1:9) %>%
  mutate(
    status = case_when(
      group == 4 & t >= 4 ~ "Treated",
      group == 5 & t >= 5 ~ "Treated",
      group == 6 & t >= 6 ~ "Treated",
      group == 8 & t >= 8 ~ "Treated",
      group == 9 & t == 9 ~ "Treated",
      group == 9 & t < 9 ~ "Last",
      TRUE ~ "Not-yet"),
    fill = case_when(
      status == "Treated" ~ "#C6E0B4",
      status == "Not-yet" ~ "#F8CBAD",
      status == "Last" ~ "#D6D6D6"),
    txt = case_when(
      status == "Treated" ~ "#385723",
      status == "Not-yet" ~ "#833C0C",
      TRUE ~ "black")) %>%
  left_join(group_df %>% dplyr::select(group, ymid), by = "group") %>%
  left_join(time_x, by = "t")

# Plot -------------------------------------------------------------------------

plot_frame <- ggplot() +

  annotate("rect", xmin = x_type_min, xmax = x_group_max, 
           ymin = y_q_min, ymax = y_year_max, fill = "white", color = "black",
           linewidth = inner_line) +
  annotate("rect", xmin = time_x$xmin[1], xmax = time_x$xmax[1], 
           ymin = y_year_min, ymax = y_year_max, fill = "white", color = "black",
           linewidth = inner_line) +
  annotate("rect", xmin = time_x$xmin[2], xmax = time_x$xmax[5], 
           ymin = y_year_min, ymax = y_year_max, fill = "white", color = "black",
           linewidth = inner_line) +
  annotate("rect", xmin = time_x$xmin[6], xmax = time_x$xmax[9],
           ymin = y_year_min, ymax = y_year_max, fill = "white", color = "black",
           linewidth = inner_line) +
  annotate("text", x = time_x$xmid[1], y = y_year_mid, label = "2015",
           family = fig_font, size = 3) +
  annotate("text", x = mean(c(time_x$xmin[2], time_x$xmax[5])), y = y_year_mid,
           label = "2016", family = fig_font, size = 3) +
  annotate("text", x = mean(c(time_x$xmin[6], time_x$xmax[9])), y = y_year_mid,
           label = "2017", family = fig_font, size = 3) +
  annotate("rect", xmin = time_x$xmin, xmax = time_x$xmax, ymin = y_q_min,
           ymax = y_q_max, fill = "white", color = "black", linewidth = inner_line) +
  annotate("text", x = time_x$xmid, y = y_q_mid,
           label = c("Q4", "Q1", "Q2", "Q3", "Q4", "Q1", "Q2", "Q3", "Q4"),
           family = fig_font, size = 3) +
  annotate("rect", xmin = time_x$xmin, xmax = time_x$xmax, ymin = y_t_min,
           ymax = y_t_max, fill = "white", color = "black", linewidth = inner_line) +
  annotate("text", x = time_x$xmid, y = y_t_mid, label = paste0("italic(t)==", 1:9),
           family = fig_font, size = 3, parse = TRUE)+
  annotate("rect", xmin = c(x_type_min, x_prov_min, x_dist_min, x_group_min),
           xmax = c(x_type_max, x_prov_max, x_dist_max, x_group_max),
           ymin = y_t_min, ymax = y_t_max, fill = "white", color = "black",
           linewidth = inner_line) +
  annotate("text", x = c(mean(c(x_type_min, x_type_max)),
                         mean(c(x_prov_min, x_prov_max)),
                         mean(c(x_dist_min, x_dist_max))),
           y = y_t_mid, label = c("Type", "Province", "District"), 
           family = fig_font, size = 3) +
  annotate("text", x = x_group_max - 0.12, y = y_t_mid, 
           label = "'Time ('*italic(t)*')'", family = fig_font, size = 3,
           parse = TRUE, hjust = 1) +
  
  # Type, separate merged columns
  annotate("rect", xmin = x_int_min, xmax = x_int_max, ymin = y_rows$ymin[4],
           ymax = y_rows$ymax[1], fill = "white", color = "black", 
           linewidth = inner_line) +
  annotate("rect", xmin = x_con_min, xmax = x_con_max, ymin = y_rows$ymin[5],
           ymax = y_rows$ymax[2], fill = "white", color = "black", 
           linewidth = inner_line) +
  annotate("text", x = mean(c(x_int_min, x_int_max)), 
           y = mean(c(y_rows$ymin[4], y_rows$ymax[1])),
           label = "Intervention", angle = 90, family = fig_font, size = 3) +
  annotate("text", x = mean(c(x_con_min, x_con_max)), 
           y = mean(c(y_rows$ymin[5], y_rows$ymax[2])), label = "Control",
           angle = 90, family = fig_font, size = 3) +

  # empty cell beside Intervention
  annotate("rect", xmin = x_con_min, xmax = x_con_max, ymin = y_rows$ymin[1],
           ymax = y_rows$ymax[1], fill = "white", color = "black",
           linewidth = inner_line) +
  
  # empty cell beside Control
  annotate("rect", xmin = x_int_min, xmax = x_int_max, ymin = y_rows$ymin[5],
           ymax = y_rows$ymax[5], fill = "white", color = "black", 
           linewidth = inner_line) +
  annotate("rect", xmin = x_prov_min, xmax = x_prov_max, ymin = y_rows$ymin,
           ymax = y_rows$ymax, fill = "white", color = "black", linewidth = inner_line) +
  annotate("rect", xmin = x_dist_min, xmax = x_dist_max, ymin = y_rows$ymin,
           ymax = y_rows$ymax, fill = "white", color = "black", linewidth = inner_line) +
  annotate("rect", xmin = x_group_min, xmax = x_group_max, ymin = y_rows$ymin,
           ymax = y_rows$ymax, fill = "white", color = "black", linewidth = inner_line) +
  annotate("text", x = mean(c(x_prov_min, x_prov_max)), y = group_df$ymid,
           label = group_df$province, family = fig_font, size = 3) +
  annotate("text", x = mean(c(x_dist_min, x_dist_max)), y = group_df$ymid,
           label = group_df$district, family = fig_font, size = 3) +
  annotate("text", x = mean(c(x_group_min, x_group_max)), y = group_df$ymid,
           label = paste0('"Group ', group_df$group, ' ("~italic(g)~"="~', 
                          group_df$group, '~")"'), 
           family = fig_font, size = 3, parse = TRUE) + 
  
  # treatment fills without borders
  annotate("rect", xmin = status_df$xmin, xmax = status_df$xmax,
           ymin = status_df$ymid - row_h / 2, ymax = status_df$ymid + row_h / 2,
           fill = status_df$fill, color = NA) +
  
  # one border around the whole treatment block
  annotate("rect", xmin = time_start, xmax = x_end, ymin = y_bottom, 
           ymax = y_body_top, fill = NA, color = "black", 
           linewidth = inner_line) +
  annotate("text", x = status_df$xmid, y = status_df$ymid, 
           label = status_df$status, color = status_df$txt, family = fig_font,
           size = 3) +
  annotate("rect", xmin = x_type_min, xmax = x_end, ymin = y_bottom,
           ymax = y_top, fill = NA, color = "black", linewidth = outer_line) +
  coord_fixed(ratio = 1, xlim = c(x_type_min, x_end), ylim = c(y_bottom, y_top),
              expand = FALSE, clip = "off") +
  theme_void() +
  theme(
    text = element_text(family = fig_font),
   # plot.margin = margin(0, 50, 0, 50)) # original -> pdf_width <- 9
   # plot.margin = margin(t = 1, r = 25, b = 1, l = 25)) -> pdf_width <- 8.5
   #plot.margin = margin(t = 1, r = 20, b = 1, l = 20)) # -> pdf_width <- 8
   plot.margin = margin(t = 1, r = 10, b = 1, l = 10)) # -> pdf_width <- 7.75

plot_frame

pdf_width <- 7.75 # increasing this also adds margin within the cell / decreasing this increases text size

pdf_height <- pdf_width * (y_top - y_bottom) / (x_end - x_type_min)
pdf("./graph/fig_02_analysis_framework.pdf",
    width = pdf_width, height = pdf_height)
print(plot_frame)
dev.off()

```
<br>

# Supplementary Figure A1. Propensity score distribution by control group
<br>
```{r fig_a_01, warning=FALSE}

# Reference
# https://cran.r-project.org/web/packages/cobalt/vignettes/cobalt.html#why-cobalt
# https://ngreifer.github.io/WeightIt/articles/WeightIt.html

# overlap 1: treated vs never-treated group
# ------------------------------------------------------------------------------

data.ovl <- data %>%
  filter(quarter == 2015.4) %>%              
  mutate(psgroup = ifelse(group > 0, 1, 0))

pscore.tr <- stats::glm(psgroup ~ logfacn + semp + urba + imps + litr,
                        family = "binomial", data = data.ovl)
ps.fit <- as.vector(pscore.tr$fitted.values)
ps.fit2 <- pmin(ps.fit, 1 - 1e-16)  # Adjust fitted values to avoid extremes

stat.ovl <- WeightIt::weightit(
  psgroup ~ logfacn + semp + urba + imps + litr,
  data = data.ovl,
  method = "glm",
  estimand = "ATT")

bal.tab(stat.ovl) %>% put()
bal.plot(stat.ovl, var.name = "prop.score", 
         which = "both", type = "histogram", mirror = TRUE)
         # "both" shows unajusted and adjusted results
bal.plot(stat.ovl, var.name = "prop.score", 
         which = "both", mirror = FALSE) # "unadjusted" only shows adjusted results


# overlap 2: treated vs last treated cohort as control
# ------------------------------------------------------------------------------

data.ovl2 <- data %>%
  filter(group > 0) %>%                         # drop group = 0
  filter(quarter == 2015.4) %>%                 # filter unique quarter
  mutate(psgroup = ifelse(group == 9, 0, 1))    # last cohort as control
        
stat.ovl2 <- WeightIt::weightit(
  psgroup ~ logfacn + semp + urba + imps + litr,
  data = data.ovl2,
  method = "glm",
  estimand = "ATT")

bal.tab(stat.ovl2) %>% put()
bal.plot(stat.ovl2, var.name = "prop.score", 
         which = "both", type = "histogram", mirror = TRUE)
bal.plot(stat.ovl2, var.name = "prop.score",
         which = "both", mirror = FALSE) # "unadjusted" only shows adjusted results


# overlap 3: treated vs last treated cohort + not-yet-treated as control
# ------------------------------------------------------------------------------

data.ovl2 <- data %>%
  filter(group > 0) %>%                         # drop group=0
  filter(quarter == 2015.4) %>%                 # filter unique quarter
  mutate(psgroup = ifelse(group == 9, 0, 1))    # last cohort as control

data.ovlnyt <- data.ovl2 %>%                    # extract
  filter(psgroup > 0) %>%                       # filter nyt group
  filter(group > 4) %>%                         # exclude the first treated
  mutate(psgroup = ifelse(psgroup == 1, 0, 0))  # nyt cohort as control

data.ovl3 <- rbind(data.ovl2, data.ovlnyt)      # combine dataset
        
stat.ovl3 <- WeightIt::weightit(
  psgroup ~ logfacn + semp + urba + imps + litr,
  data=data.ovl3,
  method = "glm",
  estimand = "ATT")

summary(stat.ovl3) %>% put()
bal.tab(stat.ovl3) %>% put()
bal.plot(stat.ovl3, var.name = "prop.score", 
         which = "both", type = "histogram", mirror = TRUE)
bal.plot(stat.ovl3, var.name = "prop.score",
         which = "both", mirror = FALSE) # "unadjusted" only shows adjusted results
         
love.plot(stat.ovl3, binary = "std", thresholds = c(m = .1))

# overlap 4: treated vs never-treated + all not-yet-treated group
# ------------------------------------------------------------------------------

data.ovlnvt <- data %>%
  filter(group == 0) %>%                   # keep never-treated group
  filter(quarter == 2015.4) %>%
  mutate(psgroup = 0)
data.ovl4 <- rbind(data.ovl3, data.ovlnvt) # combine dataset

stat.ovl4 <- WeightIt::weightit(
  psgroup ~ logfacn + semp + urba + imps + litr,
  data = data.ovl4,
  method = "glm",
  estimand = "ATT")

bal.tab(stat.ovl4) %>% put()
bal.plot(stat.ovl4, var.name = "prop.score",
         which = "both", type = "histogram", mirror = TRUE)
bal.plot(stat.ovl4, var.name = "prop.score",
         which = "both", mirror = FALSE) # "unadjusted" only shows adjusted results


# Plots ------------------------------------------------------------------------

plot.ovl1 <- cbind.data.frame(stat.ovl$ps, stat.ovl$weights, 
                              stat.ovl$treat, "nvt")
plot.ovl2 <- cbind.data.frame(stat.ovl2$ps, stat.ovl2$weights, 
                              stat.ovl2$treat, "last")
plot.ovl3 <- cbind.data.frame(stat.ovl3$ps, stat.ovl3$weights, 
                              stat.ovl3$treat, "lastnyt")

plot.ovl.all <- lapply(
 list(plot.ovl1, plot.ovl2, plot.ovl3),
 setNames, c("ps", "weight", "treat", "control_type")) %>%
 do.call(rbind, .) %>%
 mutate(
   control_type = factor(control_type,
                  levels = c("lastnyt", "last", "nvt"),
                  labels = c("Later-treated\n(not-yet- + last-treated)",
                             "Last-treated", "Never-treated")),
   group = factor(treat,
                   levels = c(0, 1),
                   labels = c("Control", "Intervention")))

write_xlsx(plot.ovl.all, "./output/fig_a_01_overlap.xlsx")

theme.ps <- function() {
 theme_bw() +
 theme(
   text = element_text(family = "Source Sans 3"), 
   panel.border = element_rect(color = "black", linewidth = 0.45),
   panel.grid.major.x = element_blank(),
   panel.grid.major.y = element_blank(),
   panel.grid.minor = element_blank(),
   panel.background = element_blank(),
   plot.margin = margin(t = 0.01, r = 0.01, b = 0.07, l = 0.04, unit = "in"),
   axis.title.x = element_text(color = "black", vjust = -1.1, size = 6.5),
   axis.title.y = element_text(color = "black", vjust = 2, size = 6.5),
   axis.text.x = element_text(color = "black", size = 6.5),
   axis.text.y = element_text(color = "black", size = 6.5),
   axis.ticks.y = element_line(color = "black", linewidth = 0.25),
   axis.ticks.x = element_line(color = "black", linewidth = 0.25),
   strip.text = element_text(color = "black", size = 6.5),
   strip.text.x = element_text(margin = margin(0, 0, 0.04, 0, "in")),
   strip.background = element_blank(),
   legend.title = element_blank(),
   legend.text = element_text(size = 6),
   legend.key.size = unit(0.15, "in"),
   legend.key.height = unit(0.05, "in"),
   legend.position = c(0.15, -0.16),
   legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "in"),
   legend.key.width = unit(0.05, "in"))
  }

# Plot with legend, scale fixed
plot.ovl.mirror.fix <- ggplot() +
 geom_histogram(data = subset(plot.ovl.all, treat == 0),
                aes(x = ps, y = after_stat(count / sum(count)),
                    fill = "Control"),
                bins = 30, col = "black", lwd = 0.2) +
 geom_histogram(data = subset(plot.ovl.all, treat == 1),
                aes(x = ps, y = -after_stat(count / sum(count)),
                    fill = "Intervention"),
                bins = 30, col = "black", lwd = 0.2) +
 facet_wrap(~control_type, ncol = 3) +
 scale_y_continuous(name = "Proportion") +
 scale_x_continuous(name = "Propensity score",
                    limits = c(-0.1, 1.1), seq(0, 1, by = 0.2)) +
 scale_fill_manual(
   values = c("Control" = "#F8766D", "Intervention" = "#9bd1a4"),
   guide = guide_legend(reverse = TRUE, nrow = 1)) +
 theme.ps()

plot.ovl.mirror.fix

# Vector file
pdf("./graph/fig_a_01_overlap.pdf", 
    width = 5, height = 2.1)
print(plot.ovl.mirror.fix)
dev.off()

```
<br>

# B6. Health service utilization trends
<br>
```{r pta, warning=FALSE}

# By intervention group/cohort
outname <- c("opo5r", "ipo5r", "opu5r", "ipu5r", "iplos", "delr")

table1_cohort <- data %>%
 filter(quarter >= 2015.4 & quarter <= 2017.3) %>%
 subset(., province != "01 Vientiane Capital")  # Drop Vientiane Capital

head(table1_cohort)

group_list <- list()
t_total_list <- list()
c_total_list <- list()
total_list <- list()

for (i in 1:length(outname)) {
 outn <- rlang::sym(outname[i])
 group_list[[i]] <- list()
 t_total_list[[i]] <- list()
 c_total_list[[i]] <- list()
 total_list[[i]] <- list()

 for (j in c(4, 5, 6, 8)) {
   group_list[[i]][[j]] <- list()
   t_total_list[[i]][[j]] <- list()
   c_total_list[[i]][[j]] <- list()

   # Intervention group
   group_t <- table1_cohort %>%
     mutate(group.new = j) %>%
     mutate(event.time = time - group.new) %>%
     filter(group == j) %>%
     group_by(group, event.time) %>%
     summarise(util.mean = mean(!!outn, na.rm = TRUE),
               util.sd = sd(!!outn, na.rm = TRUE),
               util.num = n()) %>%
     mutate(util.se = util.sd / sqrt(util.num),
            util.lower.ci = util.mean - qt(1 - (0.05 / 2), util.num - 1) * util.se,
            util.upper.ci = util.mean + qt(1 - (0.05 / 2), util.num - 1) * util.se) %>%
     mutate(intv = "Intervention")

   # Control group
   group_c <- table1_cohort %>%
     mutate(group.new = j) %>%
     mutate(event.time = time - group.new) %>%
     filter(group > j) %>%
     filter(event < 0) %>%
     group_by(event.time) %>%
     summarise(util.mean = mean(!!outn, na.rm = TRUE),
               util.sd = sd(!!outn, na.rm = TRUE),
               util.num = n()) %>%
     mutate(util.se = util.sd / sqrt(util.num),
            util.lower.ci = util.mean - qt(1 - (0.05 / 2), util.num - 1) * util.se,
            util.upper.ci = util.mean + qt(1 - (0.05 / 2), util.num - 1) * util.se) %>%
     mutate(group = j) %>%
     mutate(intv = "Control")

   group_tc <- rbind(group_t, group_c)
   group_tc$outcome <- outname[i]
   group_tc$util.sd <- NULL
   group_list[[i]][[j]] <- group_tc

   # Total
   t_total <- table1_cohort %>%
     mutate(group.new = j) %>%
     mutate(event.time = time - group.new) %>%
     filter(group == j)

   t_total_list[[i]][[j]] <- t_total

   c_total <- table1_cohort %>%
     mutate(group.new = j) %>%
     mutate(event.time = time - group.new) %>%
     filter(group > j) %>%
     filter(event < 0)

   c_total_list[[i]][[j]] <- c_total
 }

 # Combined all possible control units
 t_total_unlist <- unlist(t_total_list[i], recursive = FALSE)
 t_total_bind <- do.call("rbind", t_total_unlist)

 c_total_unlist <- unlist(c_total_list[i], recursive = FALSE)
 c_total_bind <- do.call("rbind", c_total_unlist)

 t_total_comb <- t_total_bind %>%
   group_by(event.time) %>%
   summarise(util.mean = mean(!!outn, na.rm = TRUE),
             util.sd = sd(!!outn, na.rm = TRUE),
             util.num = n()) %>%
   mutate(util.se = util.sd / sqrt(util.num),
          util.lower.ci = util.mean - qt(1 - (0.05 / 2), util.num - 1) * util.se,
          util.upper.ci = util.mean + qt(1 - (0.05 / 2), util.num - 1) * util.se) %>%
   mutate(intv = "Intervention")

 c_total_comb <- c_total_bind %>%
   group_by(event.time) %>%
   summarise(util.mean = mean(!!outn, na.rm = TRUE),
             util.sd = sd(!!outn, na.rm = TRUE),
             util.num = n()) %>%
   mutate(util.se = util.sd / sqrt(util.num),
          util.lower.ci = util.mean - qt(1 - (0.05 / 2), util.num - 1) * util.se,
          util.upper.ci = util.mean + qt(1 - (0.05 / 2), util.num - 1) * util.se) %>%
   mutate(intv = "Control")

 tc_total <- rbind(t_total_comb, c_total_comb)
 tc_total$outcome <- outname[i]
 tc_total$util.sd <- NULL
 total_list[[i]] <- tc_total
 }

# Control group trends appear similar across cohorts but differ subtly in early event.time periods. In the later post-NHI period (event time 3–4), the trends are identical because units being used are small and identical, mainly from groups 8 and 9 (last treated). Calculations are correct.

group_unlist <- unlist(group_list, recursive = FALSE)
total_bind <- do.call("rbind", total_list)
group_bind <- do.call("rbind", group_unlist)

# write_xlsx(total_bind,"./output/sfig_02_pta_qrt_total.xlsx")
# write_xlsx(group_bind,"./output/sfig_03_pta_qrt_group.xlsx")

# Total
# Averaging all units directly vs. averaging by group with weights should yield 
# identical results

CS_compute <- group_bind %>%
 group_by(event.time, intv, outcome) %>%
 mutate(weight = util.num / sum(util.num)) %>%
 mutate(avgdyn = sum(util.mean * weight)) %>%
 dplyr::select(c("event.time", "intv", "outcome", "avgdyn")) %>%
 unique() %>%
 arrange(outcome, intv, event.time)

CS_crosscheck <- left_join(CS_compute, total_bind,
                           by = c("event.time", "intv", "outcome")) %>%
  dplyr::select(contains(c("avgdyn", "util.mean"))) %>%
  print(n = Inf) %>% put()

# The results are identical
# Average of all units from each cohort by event time is equivalent to
# 1) average by cohort, and 2) average with weighting by unit size (CS code)

```

# Supplementary Figure B1. Quarterly health service utilization trends, overall
<br>
```{r fig_b_01, warning=FALSE}

var_list <- c("opo5r", "ipo5r", "opu5r", "ipu5r", "iplos", "delr")
var_label <- c("Outpatient visits per 1000, aged ≥ 5",
               "Inpatient admissions per 100 000, aged ≥ 5",
               "Outpatient visits per 1000, aged < 5",
               "Inpatient admissions per 100 000, aged < 5",
               "Inpatient average length of stay",
               "Institutional birth rate")

total_bind <- total_bind %>%
  mutate(outcome2 = factor(outcome, 
                           levels = var_list, 
                           labels = var_label))

xl <- c(min(total_bind$event.time), # x-axis length / index
          max(total_bind$event.time))

write_xlsx(total_bind,"./output/fig_b_01_pta_overall.xlsx")

plot <- ggplot(
  total_bind, aes(x = event.time, y = util.mean, 
                  group = intv, color = intv)) + 
  facet_wrap(~outcome2, ncol = 2, scales = "free") + 
  geom_vline(
    aes(xintercept = 0), linetype = "99", 
    color = "grey75", size = 0.2) +
  geom_ribbon(
    data = subset(total_bind, intv == "Intervention"),
    aes(ymin = util.lower.ci, ymax = util.upper.ci),
    fill = "#c5dec9", color = "#bad9bf", alpha = 0.65, 
    linetype = 1, linewidth = 0.025) +
  geom_ribbon(
    data = subset(total_bind, intv == "Control"),
    aes(ymin = util.lower.ci, ymax = util.upper.ci),
    fill = "#f5cdc6", color = "#f5beb5", alpha = 0.8, 
    linetype = 1, linewidth = 0.025) +
  geom_point(data = subset(total_bind, intv == "Control"), size = 0.6) +
  geom_line(data = subset(total_bind, intv == "Control"), linewidth = 0.3) +
  geom_point(data = subset(total_bind, intv == "Intervention"), size = 0.6) +
  geom_line(data = subset(total_bind, intv == "Intervention"), linewidth = 0.3) +
  scale_y_continuous(name = "Health service utilization") +
  scale_x_continuous(name = "Time since NHI introduction, quarter", 
                     breaks = xl[1]:xl[2]) +
  scale_color_manual(values = c("Control" = "#e85238", "Intervention" = "#006918"),
                     guide = guide_legend(reverse = TRUE, nrow = 1)) +
  theme_bw() +
  theme(
    text = element_text(family = "Source Sans 3"),
    panel.border = element_rect(color = "black", linewidth = 0.45),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    plot.margin = margin(t = 0.01, r = 0.05, b = 0.075, l = 0.05, unit = "in"),
    axis.title.x = element_text(color = "black", size = 6.75, 
                                margin = margin(0.125, 0, 0, 0, "in")),
    axis.title.y = element_text(color = "black", size = 6.75),
    axis.ticks = element_line(color = "black", linewidth = 0.2),
    axis.text = element_text(color = "black", size = 6.75),
    strip.text = element_text(color = "black", size = 6.75),
    strip.text.x = element_text(margin = margin(0.05, 0, 0.05, 0, "in")),
    strip.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 6.25),
    legend.key.size = unit(0.2, "in"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "in"),
    legend.position = "inside", 
    legend.position.inside =  c(0.1, -0.067))

# Vector file
pdf("./graph/fig_b_01_pretrend_overall.pdf", 
    width = 5, height = 5.4)
print(plot)
dev.off()

```

# Supplementary Figure B2. Quarterly health service utilization trends by intervention group
<br>
```{r fig_b_02, warning=FALSE}

var_list <- c("opo5r", "ipo5r", "opu5r", "ipu5r", "iplos", "delr")
var_label_l <- c("Outpatient\nvisits\nper 1000,\naged ≥ 5",
                 "Inpatient\nadmissions\nper 100 000,\naged ≥ 5",
                 "Outpatient\nvisits\nper 1000,\naged < 5",
                 "Inpatient\nadmissions\nper 100 000, \naged < 5",
                 "Inpatient\naverage\nlength\nof stay",
                 "Institutional\nbirth\nrate")

group_bind <- group_bind %>%
  mutate(cohort = factor(group, 
                         levels = c(4, 5, 6, 8),
                         labels = c("NHI introduced in \nQ3/2016\nN=15",
                                    "NHI introduced in \nQ4/2016\nN=7",
                                    "NHI introduced in \nQ1/2017\nN=37",
                                    "NHI introduced in \nQ3/2017\nN=24")),
         outcome2 = factor(outcome, levels = var_list, labels = var_label_l))

xl <- c(min(group_bind$event.time), # x-axis length / index
          max(group_bind$event.time))

write_xlsx(group_bind,"./output/fig_b_02_pta_cohort.xlsx")

plot <- ggplot(
  group_bind, aes(x = event.time, y = util.mean, group = intv, color = intv)) +
  facet_grid(outcome2 ~ cohort, scales = "free", switch = "y") +
  geom_vline(aes(xintercept = 0), linetype = "99", 
             color = "grey75", size = 0.2) +
  geom_ribbon(data = subset(group_bind, intv == "Intervention"),
              aes(x = event.time, ymin = util.lower.ci, ymax = util.upper.ci),
              fill = "#c5dec9", color = "#bad9bf", alpha = 0.65, 
              linetype = 1, linewidth = 0.025) +
  geom_ribbon(data = subset(group_bind, intv == "Control"),
              aes(x = event.time, ymin = util.lower.ci, ymax = util.upper.ci),
              fill = "#f5cdc6", color = "#f5beb5", alpha = 0.8, 
              linetype = 1, linewidth = 0.025) +
  geom_point(data = subset(group_bind, intv == "Control"), 
             aes(x = event.time, y = util.mean), size = 0.6) +
  geom_line(data = subset(group_bind, intv == "Control"), 
            aes(x = event.time, y = util.mean), linewidth = 0.3) +
  geom_point(data = subset(group_bind, intv == "Intervention"), 
             aes(x = event.time, y = util.mean), size = 0.6) +
  geom_line(data = subset(group_bind, intv == "Intervention"), 
            aes(x = event.time, y = util.mean), linewidth = 0.3) +
  scale_y_continuous(name = "Health service utilization") +
  scale_x_continuous(name = "Time since NHI introduction, quarter",
                     breaks = xl[1]:xl[2]) +
  scale_color_manual(values = c("Control" = "#e85238", 
                                "Intervention" = "#006918"),
                     guide = guide_legend(reverse = TRUE, nrow = 1)) +
  theme_bw() +
  theme(
    text = element_text(family = "Source Sans 3"),
    panel.border = element_rect(color = "black", linewidth = 0.45),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(t = 0.01, r = 0.05, b = 0.075, l = 0.05, unit = "in"),
    axis.title.x = element_text(color = "black", size = 6.75, 
                                margin = margin(0.125, 0, 0, 0, "in")),
    axis.title.y = element_blank(),
    axis.ticks.y = element_line(color = "black", linewidth = 0.2),
    axis.ticks.x = element_line(color = "black", linewidth = 0.2),
    axis.text.y = element_text(color = "black", size = 6.75),
    axis.text.x = element_text(color = "black", size = 6.75),
    strip.placement = "outside",
    strip.text = element_text(color = "black", size = 6.75),
    strip.text.y.left = element_text(
      angle = 0, hjust = 0, 
      margin = margin(t = 0, r = 0, b = 0, l = 0, "in")),
    strip.text.x = element_text(margin = margin(0, 0, 0.05, 0, "in")),
    strip.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 6.25),
    legend.key.size = unit(0.2, "in"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "in"),
    legend.position = "inside", 
    legend.position.inside =  c(0.05, -0.057))

# Vector file
pdf("./graph/fig_b_02_pretrend_cohort.pdf", 
    width = 5, height = 6.25)
print(plot)
dev.off()

```

# Log close
<br>
```{r log_close, warning=FALSE}

# Log close
log_close(footer = TRUE)
writeLines(readLines(logrecord))

```
<br>